\input{Docs/Paper/Sections/general settings}





 


 \title{ Gender Composition in Classrooms: Influences on Post-Secondary Schooling  Choices \thanks{I thank ....
} }

 
\author{
{ 
\href{https://polanco-jaime.github.io/}{Jaime Polanco-Jiménez}\thanks{Corresponding author, PhD Student at KU Leuven \& Pontificia Universidad Javeriana.    (email:
\href{mailto:jaime.polanco@javeriana.edu.co}{jaime.polanco@javeriana.edu.co} ) %, Tel: +57 3002737490.
  }   
 }\\
  {
Kristof De Witte \thanks{ KU Leuven \& UNU-Merit, Maastricht University.  (email: \href{mailto:: kristof.dewitte@kuleuven.be}{ kristof.dewitte@kuleuven.be}).  } 
 }
 \\
  {
Gloria L. Bernal \thanks{  Pontificia Universidad Javeriana.  (email: \href{mailto:: gbernal@javeriana.edu.co}{ gbernal@javeriana.edu.co}).  } 
 }   
}

 
 \date{\normalsize\today} 
% \singlespacing
\doublespacing 
% \onehalfspacing

\begin{document}
\renewcommand{\BOthers}[1]{et al.\hbox{}}
\maketitle

\abstract{ 
\singlespacing 
% \onehalfspacing
\noindent   
 
This study explores the impact of male students' presence in classrooms on the choice of university majors among female students. To unpack this, we employ a fixed-effects methodology and a staggered difference-in-difference (S-DiD) approach, specifically examining gender compositions within classrooms concerning university major choices.Our findings reveal that, compared to their male counterparts, female students show a significantly higher inclination toward pursuing academic majors associated with human sciences. These results suggest that classrooms with a high composition of male students narrow the gap in STEM majors for women, except in fields such as Law and Medicine. Furthermore, no gender differences were found based on classroom gender composition in majors related to agriculture, zootechnics, and veterinary sciences.


\vspace*{0.3cm}
\textit{JEL} Codes:   O28, I21.\\ %, L71, C93, C21.
\vspace*{0.1cm}
% \vspace*{0.3cm} 
\textbf{Keywords: } 
\\
}

\vfill

%---------------------------------------------------------------------------------------
\newpage    

         
 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{section:01}
 

\input{Docs/Paper/Sections/01 Section}
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Context and Background}
\section{Data and Descriptive Statistics}\label{section:02}
\input{Docs/Paper/Sections/02 Section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ Empirical Strategy}\label{section:03}
\input{Docs/Paper/Sections/03 Section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results} \label{section:04}
\input{Docs/Paper/Sections/04 Section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions} \label{section:05}
\input{Docs/Paper/Sections/05 Section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  
 \newpage
\bibliographystyle{apacite}

\bibliography{Rereferences/bibliography}

\appendix 
\newpage
\input{Docs/Paper/Sections/06 Appendix}\label{section:app}
\end{document}




 
